---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# managelidar

<!-- badges: start -->

<!-- badges: end -->

The goal of managelidar is to facilitate the handling and management of lidar data files (`*.laz`), e.g. to provide some functions to get a quick overview of file attributes. This should facilitate the quality check of incoming ALS data. 
`managelidar` makes use of functions provided by {lidR} and {lasR}. Functions are not reading the entire LAZ files, as this would require long computations for large collections of data. Instead, attributes are read from the file header.  

## Installation

You can install the development version of managelidar from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("wiesehahn/managelidar")
```

As some functions depend on the [lasR](https://github.com/r-lidar/lasR) package (version >= 0.7.2) which is hosted on at https://r-lidar.r-universe.dev/lasR you have to manually install it in advance with:

``` r
# Install lasR in R:
install.packages("lasR", repos = c("https://r-lidar.r-universe.dev", "https://cran.r-project.org"))
```

## Example

This is a basic function which queries the spatial extents of all lidar data files (`*.laz`) in a given folder from its data headers (without reading the actual point cloud data). It returns a dataframe which can be used in further data management steps. 

```{r get_extent}
library(managelidar)
f <- system.file("extdata", package="managelidar")
get_extent(f)
```


