---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# managelidar

<!-- badges: start -->

<!-- badges: end -->

The goal of managelidar is to facilitate the handling and management of lidar data files (`*.laz`).

## Installation

You can install the development version of managelidar from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("wiesehahn/managelidar")
```

## Example

This is a basic function which queries the spatial extents of all lidar data files (`*.laz`) in a given folder from its data headers (without reading the actual point cloud data). It returns a dataframe which can be used in further data management steps. 

```{r get_extent}
library(managelidar)
f <- system.file("extdata", package="managelidar")
get_extent(f)
```


