---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# managelidar

<!-- badges: start -->

<!-- badges: end -->

The goal of managelidar is to facilitate the handling and management of lidar data files (`*.las/laz`). 
Its's main purpose is to convert new incoming data to data with certain quality standards. Further, it provides functions to facilitate the quality check of incoming ALS data. 
`managelidar` makes use of functions provided by {lidR} and {lasR}. Most functions which provide information about 3D point cloud files are working without reading the entire files, as this would require long computations for large collections of data. Instead, attributes are read from the file header where possible.  

## Installation

You can install the development version of managelidar from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pkg_install("wiesehahn/managelidar")
```

As some functions depend on the [lasR](https://github.com/r-lidar/lasR) package (version >= 0.14.1) which is hosted on at https://r-lidar.r-universe.dev/lasR, you have to manually install it in advance with:

``` r
# Install lasR in R:
install.packages("lasR", repos = c("https://r-lidar.r-universe.dev", "https://cran.r-project.org"))
```

## Example

This is a basic function which queries the spatial extents of all lidar data files in a given folder from its data headers (without reading the actual point cloud data). It returns a dataframe which can be used in further data management steps. 

```{r get_extent}
library(managelidar)
f <- system.file("extdata", package="managelidar")
get_extent(f)
```


